---
title: "Lab 03 - Regular Expressions"
author: "Gregorio Salcedo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(dplyr)
library(stringr)
library(knitr)
library(pander)

```

### Preview Data


```{r}
URL <- "https://raw.githubusercontent.com/DS4PS/cpp-527-fall-2020/master/labs/data/medium-data-utf8-v2.csv"
d <- read.csv( URL )

preview.these <- c("title", "subtitle", "claps", "reading_time", "publication", "date")
head( d[preview.these] ) %>% pander()
```

### Pre-Processing Steps



```{r}
# Remove extra spaces, HTML tags, and hair space dashes
d$title <- gsub("\\s+", " ", d$title) 
d$title <- gsub("<[^>]+>", "", d$title) 
d$title <- gsub("\u200A—\u200A", "", d$title) 
head(d$title)

```
### Performance Measure


```{r}
hist( d$claps, main="Raw Clap Count",
      xlim=c(1,950), breaks=5000, col="gray20", border="white" )
```



```{r}
hist( log10(d$clap+1), main="Logged Clap Score",
      col="gray20", border="white", breaks=100 )
```



```{r}
d$clap.score <- log10( d$claps + 1 )

```
### Q1

Q1-A:What is the average performance (clap score) of each type of title?
  How-to Guide        2.189      

   Power List         2.166      

    Question          2.043      

     Other            2.039      

  Colon Title         1.964  
  
Q1-B:Which style of title performs the best overall?
How-to Guide	
```{r}
# Categorize titles 
d$title_category <- case_when(
  grepl("^[0-9]+\\s", d$title) ~ "Power List", 
  grepl("^(How to|How we)", d$title, ignore.case = TRUE) ~ "How-to Guide",
  grepl(":.*", d$title) ~ "Colon Title",  # Contains a colon
  grepl("\\?$", d$title) | grepl("^(Why|What|How|Can)\\s", d$title, ignore.case = TRUE) ~ "Question",
  TRUE ~ "Other"  
)


head(d[, c("title", "title_category")]) %>% pander()
```


```{r}
# average clap score
avg_clap_scores <- d %>%
  group_by(title_category) %>%
  summarise(avg_clap_score = mean(clap.score, na.rm = TRUE)) %>%
  arrange(desc(avg_clap_score))

avg_clap_scores %>% pander()

```


```{r}
best_category <- avg_clap_scores %>% slice(1)
best_category
```
### Q2

Q2-A: What are the 25 most common words used in the titles?
  to    the     a     how   of    and   your   in    for   you   is    with 
   why   data   what    i    on    from   an    learning   can   are   my    be    &nbsp; 

Q2-B: What is the most common word used at the beginning of a title?
How
Q2-C: What is the most common word used at the end of a title?
Machine Learning

```{r}
#lowercase 
d$title_clean <- tolower(d$title)

#split titles
word_list <- strsplit(d$title_clean, "\\s+")

head (word_list)

```


```{r}
#get all words in one vector 
all_words <- unlist(word_list)

# Remove punctuation
all_words <- gsub("[[:punct:]]", "", all_words)

# Count
word_freq <- table(all_words)

#top 25
top_25_words <- sort(word_freq, decreasing = TRUE)[1:25]

top_25_words %>% pander()

```


```{r}
# extract first words
first_words <- sapply(word_list, function(x) x[1])

# count first word frequency
first_word_freq <- table(first_words)

# most common first word
most_common_first <- names(sort(first_word_freq, decreasing = TRUE))[1]

most_common_first
```



```{r}
# Extract last words
last_words <- sapply(word_list, function(x) x[length(x)])

# Count last word frequency
last_word_freq <- table(last_words)

# most common last word
most_common_last <- names(sort(last_word_freq, decreasing = TRUE))[1]

most_common_last
```



